Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-19 TACAS2018/paper.tex
authors

We are interested in control systems with an output $\vec{y}$ that is a linear combination of its states and inputs:
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t).
\end{equation}
%
%In this work, 
Equations~\eqref{eq:ode1}-\eqref{eq:ode2} are soundly discretized in time~\cite{middleton1990digital,van1978computing} \note{mention sample time, referred to later.}
(cf. Appendix~\ref{sec:cont_aa}) into
\begin{align}
\label{eq:plant}
\left\{ 
\begin{array}{l}
\vec{x}_{k+1} = \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} = \mat{C}_d \vec{x}_k + \mat{D}_d \vec{u}_k, 
\end{array}
\right.
\end{align} 
%
where $k \in \mathbb N$ is a discrete counter and $x_{0}=x(0)$ is the initial state. 
Matrices $\mat{A}_d$, $\mat{B}_d$, $\mat{C}_d$ and $\mat{D}_d$ are equal-sized to $\mat{A}$, $\mat{B}$, $\mat{C}$ and $\mat{D}$ respectively,  
and charactirize the discretized plant dynamics. 
%, whereas the latter denote the continuous evolution.  
In this work we make the standard assumption that $\mat{D}_d=\mat{D}=\mat{0}$ \cite{Astrom08}.  
\note{Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.}
%% We synthesize for requirements over this discrete-time domain. 
%% Later, we will address the issue of variable quantization, 
%% as introduced by the ADC/DAC conversion blocks (Fig.~\ref{fig:digitalsystem}).

@@

The time discretization introduced in Sec.~\ref{sec:ssrepresentation} leads to evaluate the state of the model at integer multiples of the sample time. 
 However, in order to synthesise $\mat{K}$ and $\mat{L}$, 
 we must be able to ensure that certain properties of the system hold at any given (real) time. 
 Consequently, we accelerate the dynamics of the system according to the model semantics in Sec.~\ref{sec:model_semantics} in order to compute the state $\vec{x}_T$  at a given time $T$ \note{unclear: we did not say what ``accelerate'' means, only mention we compute the state of the discrete-time model at time step T.}.

@@

 \begin{theorem}
 Given $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$, we have
\note{unclear: next notation refers to discrete-time model}
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x}_0\\
 \text{where }\mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(1)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn}\\
 &\text{and } s_i=\left\{\begin{array}{cc}1&gm(\lambda_i)>1\\0&gm(\lambda_i)=1\end{array}\right.,\nonumber
 \end{align}
$\lambda_i \in \mat{J}$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$ \cite{Astrom08}.  
We call $\vec{x}_T$ a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time t=T.
 \end{theorem}
 %% We can apply Th.~\ref{eq:continuous_tube_dyn} to discretized systems without inputs in order to
 %% obtain a witness ..
 %
 \note{proof?}  
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete-dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_r=A_d^r\vec{x}_0 : r \in \mathbb{R}^+$ is a witness of the system at time $t=rT_d$.
 \end{corollary}
  \note{proof?} 
%
 % \begin{proof}
% Let us recall equation \ref{eq:discretize}. The discrete representation of the system dynamics discretised with sample time $T_1$ is ruled by the formula
% $\mat{A}_1 = e^{\mat{A} T_1}$. From matrix theory, we have 
%\begin{align}
% e^{\mat{A}}&=\sum_{k=0}^\infty \frac{1}{k!}\mat{A}^k
%\end{align} 
%\begin{align} 
% e^{\mat{S}\mat{J}\mat{S}^{-1}}&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{S}\mat{J}\mat{S}^{-1}\right)^k
% =\mat{S} \left (\sum_{k=0}^\infty \frac{1}{k!}\mat{J}^k\right) \mat{S}^{-1}\nonumber\\
% &=\mat{S}e^{\mat{J}}\mat{S}^{-1}
%\end{align} 
%\begin{equation}
%\mat{A}=\mat{S}\mat{J}\mat{S}^{-1} \Rightarrow \vec{A}_1 = \mat{S}\mat{J}_1\mat{S}^{-1}= \mat{S}e^{\mat{J} T_1}\mat{S}^{-1}.
% \end{equation}
% 
% Let $\mat{J}$ be a diagonal matrix, such that 
% $${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda_i \in \mat{J}.$$
% Let us now take a different sample rate $T_2$, such that 
% $${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda_i \in \mat{J}.$$
% We can then say that 
% \begin{equation}
% {\lambda_2}_i=e^{\lambda_i T_1 \frac{T_2}{T_1}}={\lambda_1}_i^{\frac{T_2}{T_1}} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}.
% \end{equation}
% This proves the theorem for $gm(\lambda_i)=1$.
% We now note that given an exponentiated Jordan form with geometric multiplicity, the upper diagonal terms can be
% referenced to the original eigenvalues modified by the sampling rate.
%  Let $\mat{J}\in \mathbb{R}^{p \times p}$ be a Jordan Canonical matrix.
% Then
%\begin{align}
% \mat{J}_1&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{J}T_1\right)^k=\sum_{k=0}^\infty \frac{1}{k!} \left [ \begin{array}{cccc}
% \lambda_i^k  & \binom{k}{1}  \lambda^{k-1} & \hdots  & \binom{k}{p-1} \lambda_i^{k-p+1} \\
%& \lambda_i^k  & \binom{k}{1}  \lambda_i^{k-1} & \vdots \\
%\vdots & \ddots & \ddots & \vdots \\
%& &  &\lambda_i^k \\
%\end{array} \right ] T_1^k
%\end{align}
%The eigenvalues remain the same as in the diagonal case, but the upper triangular terms are of the form:
%\begin{align}
%\forall j>i, c_{ij}&=\sum_{k=j-i}^\infty \frac{1}{k!}\binom{k}{j-i} \lambda_i^{k-(j-i)}T_1^k\\
%&=\frac{1}{(j-i)!}\sum_{k=j-i}^\infty \frac{1}{(k-(j-i))!} \lambda_i^{k-(j-i)}T_1^k\nonumber\\
%&=\frac{T_1^{j-i}e^{\lambda_i T_1}}{(j-i)!}\nonumber
%\end{align}
%which completes the proof for $gm(\lambda_i)>1$.
%\end{proof}

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics for systems with parametric inputs}\label{sec:real_discrete_param_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 Next, we are interested in observing systems with parametric
 inputs, described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$.
 For such systems, a witness $\vec{x}_T$ is computed next.
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}^{-1}\vec{x}_T'\nonumber\\
\vec{x}_T'&=\mat{A}\mat{A}_T\vec{x}_0 + (\mat{I}-\mat{A}_T)\mat{B}\vec{u} : \forall t \leq T,\ \vec{u}(t)=\vec{u} 
 \end{align}
 with $\mat{A}_T$ as per \eqref{eq:continuous_tube_dyn} is a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a parametric input $\vec{u}$.
\end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d :  A_d=e^{\mat{A} T_d}$, then $\vec{x}_r=A_d^r\vec{x}_0+(\mat{I}-\mat{A}_d^r\mat{B}\vec{u}) : r \in \mathbb{R}^+$ is a witness of the system at time $rT_d$.
 \end{corollary}
% \begin{proof}
% We begin by expanding the equation
% \begin{align}
% \vec{x}_T&=\mat{A}^{-1}\vec{x}_T'\nonumber\\
% &=\mat{A}_T\vec{x}_0 + \mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}\vec{u}\nonumber\\
% &= \mat{A}_T\vec{x}_0 + \mat{B}_T\vec{u}
% \end{align}
% where $\mat{A}_T=e^{\mat{A}T}$ and $\mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}$ correspond to $\mat{A}_d$ and $\mat{B}_d$ in equation \eqref{eq:discretize}
% \end{proof}

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics for systems with discrete time feedback inputs}\label{sec:real_discrete_feedback_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

The final case we analyse is that of hybrid time systems, where the
plant dynamics evolve in continuous time while the feedback dynamics evolve
in discrete time, i.e. a new input value is only produced for every sample interval.

\begin{theorem}
The expression
 \begin{align}
 \vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0
 \label{eq:cyber_feedback}
 \end{align}
 with $\mat{A}_T$ as per \eqref{eq:continuous_tube_dyn}, 
 $\mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}$ and 
 $\mat{A}_d, \mat{B}_d$ describing the discretized plant dynamics is a witness of the system
 \begin{align}
 \dot{\vec{x}}(t) &= \mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)\nonumber\\
 \vec{u}(t)&=\mat{K}\vec{x}_k  \mid  kT_s \leq t < (k+1)T_s
 \end{align}
 \end{theorem}
%
%
%\begin{proof}
%Let $\mat{K}$ be a feedback controller such that $\vec{u}_k=-\mat{K}\vec{x}_k$ at time $t=kT_s$. Let the output of
%the feedback system be a zero order holder such that $\vec{u}(t+kT_s)=\vec{u}(kT_s) : 0 \leq t < T_s$.
%The closed loop dynamics of the system between times $kT_s$ and $kT_s+T_s$ are
%\begin{equation}
%\vec{x}_{k,t}=\mat{A}_t\vec{x}_k-\mat{B}_t\mat{K}\vec{x}_k = (\mat{A}_t-\mat{B}_t\mat{K})\vec{x}_k : 0 \leq t < T_s
%\end{equation}
%We further explore the dynamics across discrete time boundaries. Let $\mat{A}_d=\mat{A}_{T_s}$ and $\mat{B}_d=\mat{B}_{T_s}$
%\begin{align}
%\vec{x}_{k}&=\vec{x}_{k-1,T_s}= (\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_{k-1}\\
%\vec{x}_{k,T_s} &= (\mat{A}_d-\mat{B}_d\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_{k-1}
%\end{align}
%which accelerating results becomes
%\begin{align}
%\label{eq:feedback_sampled_cont}
%\vec{x}_{k} &= (\mat{A}_d-\mat{B}_d\mat{K}) ^k\vec{x}_0\\
%\vec{x}_{t} &= (\mat{A}_{t-kT_s}-\mat{B}_{t-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0
%\label{eq:feedback_cont}
%\end{align}
%\end{proof}

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{Error modelling}\label{sec:errors}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Time delay on discrete time feedback systems} \label{sec:delay}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
A common ocurrence in digital systems is that they do not necessarily
satisfy the real-time specifications.  Delays due to processing time,
communication lines, or other mechanisms will cause disturbances in the
control loop.  In our model these delays are all deferred to the ZOH, i.e.,
we presume that everything that happens in the digital world takes place at
the sampling instance $t$ and the output of the result is delayed at the DAC
when it is fed back into the real world.  We~will also presume that this
delay is much smaller than the sampling time.

\begin{theorem}
The error introduced by a delay $t_d$ in the control output $\vec{u}_k$ of a discrete feedback controller for a continuous LTI system can be modelled as a non-deterministic noise signal upperbound by 
\begin{align}
\norm{\vec{y}_{k}-{\vec{y}_{t_d}}_{k}}_1 \leq q_t, \text{ where } q_t=&(\norm{\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})}_1+\norm{\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}}_1)\norm{U}_1
\end{align}
where $\vec{y}_{t_d}$ is the output of the system with delay $t_d$, and $\norm{\vec{x}}_1=\sum_{i=0}^{size(\vec{x})} |\vec{x}_i|$ is the $l_1$ norm (in for sets, $\norm{X}_1 = sup \norm{\vec{x}}_1 \mid \vec{x} \in X$). 
\end{theorem}
%
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Errors generated by the numerical representation} 
\label{sec:numeric_rep}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

Let us first consider the effect of the quantization errors. 
Within the controller, state values are manipulated at low precision.
Furthermore, the quantizers for the DAC and the ADC introduce truncation errors
which we denote by the effect of the functions $\mathcal{F}_{\langle D \rangle}$
and $\mathcal{F}_{\langle A \rangle}$, respectively. Finally, let us call $\vec{\nu}_e$ the
error caused by the FWL effects on the observer
\todo{the observer was not defined..}
(\emph{i.e} the numerical errors 
in the calculations as opposed to the estimation error).
The inputs are computed using the following equation: 
%
\begin{align}
u_{k}&=\mathcal{F}_{\langle D \rangle}\left(-\mat{K}_{\langle C \rangle}\cdot\left(\mathcal{F}_{\langle C \rangle}\left(\mathcal{F}_{\langle A \rangle}(\vec{x}_{k})\right)\right) +\vec{\nu}_e\right)
\label{eq:uk}
\end{align}

This induces two types of the errors: first, the truncation
error due to the domain transformations $\mathcal{F}_{\langle *
\rangle}(\cdot)$; and second, the rounding error introduced by the
multiplication operation.  We represent these errors as non-deterministic
additive noise. More details can be found in Appendix~\ref{appendix:quantization}.

@@

Previous studies~\cite{gangli1} show that the FWL affects the poles and
zeros positions, degrading the closed-loop dynamics, causing steady-state
errors and eventually de-stabilizing the system~\cite{Bessa16}.  However,
since the verifier also checks for the stability of the dynamics as a
precursor to safety, this case is handled inside the loop.  In the
following, we shall disregard these steady-state errors (caused by FWL
effects) when stability is ensured by synthesis, and then verify its safety
accounting for the finite-precision errors.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{Properties of a controllable observed system}\label{sec:cof_verification}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%% In most real implementations, the internal state of a plant cannot be
%% directly measured but rather inferred from its
%% effect on the output. A plant model that minimises the difference
%% between the outputs is called an observer and it is a well known
%% paradigm in the control literature~\cite{astrom1997computer}.

>>>>

We are interested in control systems with an output $\vec{y}$ that is a linear combination of its states and inputs:
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t).
\end{equation}
%
%In this work, 
\note{Observability requirement? Used later.}
Equations~\eqref{eq:ode1}-\eqref{eq:ode2} are soundly discretized in time~\cite{middleton1990digital,van1978computing} \note{mention sample time, referred to later.}
(cf. Appendix~\ref{sec:cont_aa}) into
\begin{align}
\label{eq:plant}
\left\{ 
\begin{array}{l}
\vec{x}_{k+1} = \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} = \mat{C}_d \vec{x}_k + \mat{D}_d \vec{u}_k, 
\end{array}
\right.
\end{align} 
%
where $k \in \mathbb N$ is a discrete counter and $x_{0}=x(0)$ is the initial state. 
Matrices $\mat{A}_d$, $\mat{B}_d$, $\mat{C}_d$ and $\mat{D}_d$ are equal-sized to $\mat{A}$, $\mat{B}$, $\mat{C}$ and $\mat{D}$ respectively,  
and charactirize the discretized plant dynamics. 
%, whereas the latter denote the continuous evolution.  
In this work we make the standard assumption that $\mat{D}_d=\mat{D}=\mat{0}$ \cite{Astrom08}.  
\note{Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.}
%% We synthesize for requirements over this discrete-time domain. 
%% Later, we will address the issue of variable quantization, 
%% as introduced by the ADC/DAC conversion blocks (Fig.~\ref{fig:digitalsystem}).

@@

The time discretization introduced in Sec.~\ref{sec:ssrepresentation} leads to evaluate the state of the model at integer multiples of the sample time. 
 However, in order to synthesise $\mat{K}$ and $\mat{L}$, 
 we must be able to ensure that certain properties of the system hold at any given (real) time. 
 Consequently, we accelerate the dynamics of the system according to the model semantics in Sec.~\ref{sec:model_semantics} in order to compute the state $\vec{x}_T$  at a given time $T$ \note{unclear: we did not say what ``accelerate'' means, you only mention that we compute the state of the discrete-time (according to your notation) model at time step T.}.

@@

 \begin{theorem}
 Given $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$, we have
\note{unclear: next notation refers to discrete-time model}
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x}_0\\
 \text{where }\mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(1)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn}\\
 &\text{and } s_i=\left\{\begin{array}{cc}1&gm(\lambda_i)>1\\0&gm(\lambda_i)=1\end{array}\right.,\nonumber
 \end{align}
$\lambda_i \in \mat{J}$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$ \cite{Astrom08}.  
We call $\vec{x}_T$ a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time t=T. 
\note{What exactly is a ``witness''? Do you mean ``the solution at that time''?}
 \end{theorem}
 %% We can apply Th.~\ref{eq:continuous_tube_dyn} to discretized systems without inputs in order to
 %% obtain a witness ..
 %
 \note{proof? So what is the use of this theorem? How does it help to address the issue discussed at the beginning of the section?}  
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete-dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_r=A_d^r\vec{x}_0 : r \in \mathbb{R}^+$ is a witness of the system at time $t=rT_d$.
 \end{corollary}
  \note{proof?} 
%
 % \begin{proof}
% Let us recall equation \ref{eq:discretize}. The discrete representation of the system dynamics discretised with sample time $T_1$ is ruled by the formula
% $\mat{A}_1 = e^{\mat{A} T_1}$. From matrix theory, we have 
%\begin{align}
% e^{\mat{A}}&=\sum_{k=0}^\infty \frac{1}{k!}\mat{A}^k
%\end{align} 
%\begin{align} 
% e^{\mat{S}\mat{J}\mat{S}^{-1}}&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{S}\mat{J}\mat{S}^{-1}\right)^k
% =\mat{S} \left (\sum_{k=0}^\infty \frac{1}{k!}\mat{J}^k\right) \mat{S}^{-1}\nonumber\\
% &=\mat{S}e^{\mat{J}}\mat{S}^{-1}
%\end{align} 
%\begin{equation}
%\mat{A}=\mat{S}\mat{J}\mat{S}^{-1} \Rightarrow \vec{A}_1 = \mat{S}\mat{J}_1\mat{S}^{-1}= \mat{S}e^{\mat{J} T_1}\mat{S}^{-1}.
% \end{equation}
% 
% Let $\mat{J}$ be a diagonal matrix, such that 
% $${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda_i \in \mat{J}.$$
% Let us now take a different sample rate $T_2$, such that 
% $${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda_i \in \mat{J}.$$
% We can then say that 
% \begin{equation}
% {\lambda_2}_i=e^{\lambda_i T_1 \frac{T_2}{T_1}}={\lambda_1}_i^{\frac{T_2}{T_1}} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}.
% \end{equation}
% This proves the theorem for $gm(\lambda_i)=1$.
% We now note that given an exponentiated Jordan form with geometric multiplicity, the upper diagonal terms can be
% referenced to the original eigenvalues modified by the sampling rate.
%  Let $\mat{J}\in \mathbb{R}^{p \times p}$ be a Jordan Canonical matrix.
% Then
%\begin{align}
% \mat{J}_1&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{J}T_1\right)^k=\sum_{k=0}^\infty \frac{1}{k!} \left [ \begin{array}{cccc}
% \lambda_i^k  & \binom{k}{1}  \lambda^{k-1} & \hdots  & \binom{k}{p-1} \lambda_i^{k-p+1} \\
%& \lambda_i^k  & \binom{k}{1}  \lambda_i^{k-1} & \vdots \\
%\vdots & \ddots & \ddots & \vdots \\
%& &  &\lambda_i^k \\
%\end{array} \right ] T_1^k
%\end{align}
%The eigenvalues remain the same as in the diagonal case, but the upper triangular terms are of the form:
%\begin{align}
%\forall j>i, c_{ij}&=\sum_{k=j-i}^\infty \frac{1}{k!}\binom{k}{j-i} \lambda_i^{k-(j-i)}T_1^k\\
%&=\frac{1}{(j-i)!}\sum_{k=j-i}^\infty \frac{1}{(k-(j-i))!} \lambda_i^{k-(j-i)}T_1^k\nonumber\\
%&=\frac{T_1^{j-i}e^{\lambda_i T_1}}{(j-i)!}\nonumber
%\end{align}
%which completes the proof for $gm(\lambda_i)>1$.
%\end{proof}

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics for models with parametric inputs}\label{sec:real_discrete_param_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 Next, we are interested in observing systems with parametric
 inputs, described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ \note{confusing: if the input is parametric, it should not depend on time}. 
 For such systems, a witness $\vec{x}_T$ is computed next.
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}^{-1}\vec{x}_T'\nonumber\\
\vec{x}_T'&=\mat{A}\mat{A}_T\vec{x}_0 + (\mat{I}-\mat{A}_T)\mat{B}\vec{u} : \forall t \leq T,\ \vec{u}(t)=\vec{u} 
 \end{align}
 with $\mat{A}_T$ as per \eqref{eq:continuous_tube_dyn} is a witness of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a parametric input $\vec{u}$. 
 \note{Just write u, w/o time dependence.}
\end{theorem}
 \begin{corollary}
 If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d :  A_d=e^{\mat{A} T_d}$, then $\vec{x}_r=A_d^r\vec{x}_0+(\mat{I}-\mat{A}_d^r\mat{B}\vec{u}) : r \in \mathbb{R}^+$ is a witness of the system at time $rT_d$.
 \end{corollary}
% \begin{proof}
% We begin by expanding the equation
% \begin{align}
% \vec{x}_T&=\mat{A}^{-1}\vec{x}_T'\nonumber\\
% &=\mat{A}_T\vec{x}_0 + \mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}\vec{u}\nonumber\\
% &= \mat{A}_T\vec{x}_0 + \mat{B}_T\vec{u}
% \end{align}
% where $\mat{A}_T=e^{\mat{A}T}$ and $\mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}$ correspond to $\mat{A}_d$ and $\mat{B}_d$ in equation \eqref{eq:discretize}
% \end{proof}

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated dynamics for models with discrete-time feedback inputs}\label{sec:real_discrete_feedback_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

The final case we analyse is that of hybrid-time models, 
where the plant dynamics evolve in continuous time while the feedback dynamics evolve in discrete time, 
namely a new input value is produced only at every sample time.
%
\begin{theorem}
The expression
 \begin{align}
 \vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0, 
 \label{eq:cyber_feedback}
 \end{align}
 with $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn}, 
 $\mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B}$ and 
 $\mat{A}_d, \mat{B}_d$ describing the discretized plant dynamics is a witness of the system
 \begin{align}
 \dot{\vec{x}}(t) = \mat{A}\vec{x}(t)+\mat{B}\vec{u}(t), \quad 
 \vec{u}(t)=\mat{K}\vec{x}_k,  
 kT_s \leq t < (k+1)T_s. 
 \end{align}
 \end{theorem}
%
%
%\begin{proof}
%Let $\mat{K}$ be a feedback controller such that $\vec{u}_k=-\mat{K}\vec{x}_k$ at time $t=kT_s$. Let the output of
%the feedback system be a zero order holder such that $\vec{u}(t+kT_s)=\vec{u}(kT_s) : 0 \leq t < T_s$.
%The closed loop dynamics of the system between times $kT_s$ and $kT_s+T_s$ are
%\begin{equation}
%\vec{x}_{k,t}=\mat{A}_t\vec{x}_k-\mat{B}_t\mat{K}\vec{x}_k = (\mat{A}_t-\mat{B}_t\mat{K})\vec{x}_k : 0 \leq t < T_s
%\end{equation}
%We further explore the dynamics across discrete time boundaries. Let $\mat{A}_d=\mat{A}_{T_s}$ and $\mat{B}_d=\mat{B}_{T_s}$
%\begin{align}
%\vec{x}_{k}&=\vec{x}_{k-1,T_s}= (\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_{k-1}\\
%\vec{x}_{k,T_s} &= (\mat{A}_d-\mat{B}_d\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_{k-1}
%\end{align}
%which accelerating results becomes
%\begin{align}
%\label{eq:feedback_sampled_cont}
%\vec{x}_{k} &= (\mat{A}_d-\mat{B}_d\mat{K}) ^k\vec{x}_0\\
%\vec{x}_{t} &= (\mat{A}_{t-kT_s}-\mat{B}_{t-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0
%\label{eq:feedback_cont}
%\end{align}
%\end{proof}

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{Modelling Errors in the Control Loop}\label{sec:errors}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Time delay on discrete-time feedback models} \label{sec:delay}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

A common ocurrence in digital systems is that they do not necessarily
satisfy the real-time specifications.  Delays due to processing time,
communication lines, or other mechanisms will cause disturbances in the
control loop.  In our model these delays are all deferred to the ZOH, i.e.,
we presume that everything that happens in the digital world takes place at
the sampling instance $t$ and the output of the result is delayed at the DAC
when it is fed back into the real world.  We~will also presume that this
delay is much smaller than the sampling time.

\begin{theorem}
The error introduced by a delay $t_d$ in the control output $\vec{u}_k$ of a discrete feedback controller for a continuous LTI system can be modelled as a non-deterministic noise signal upperbound by 
\begin{align}
\norm{\vec{y}_{k}-{\vec{y}_{t_d}}_{k}}_1 \leq q_t, \text{ where } q_t=&(\norm{\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})}_1+\norm{\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}}_1)\norm{U}_1
\end{align}
where $\vec{y}_{t_d}$ is the output of the system with delay $t_d$, and $\norm{\vec{x}}_1=\sum_{i=0}^{size(\vec{x})} |\vec{x}_i|$ is the $l_1$ norm (in for sets, $\norm{X}_1 = \sup \norm{\vec{x}}_1, \vec{x} \in X$). 
\end{theorem}
%
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Errors generated by numerical representations} 
\label{sec:numeric_rep}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

Let us first consider the effect of the quantization errors. 
Within the controller, state values are manipulated at low precision.
Furthermore, the quantizers for the DAC and the ADC introduce truncation errors
which we denote by the effect of the functions $\mathcal{F}_{\langle D \rangle}$
and $\mathcal{F}_{\langle A \rangle}$, respectively. Finally, let us call $\vec{\nu}_e$ the
error caused by the FWL effects on the observer
\todo{the observer was not defined..}
(\emph{i.e} the numerical errors 
in the calculations as opposed to the estimation error).
The inputs are computed using the following equation: 
%
\begin{align}
u_{k} =\mathcal{F}_{\langle D \rangle}\left(-\mat{K}_{\langle C \rangle}\cdot\left(\mathcal{F}_{\langle C \rangle}\left(\mathcal{F}_{\langle A \rangle}(\vec{x}_{k})\right)\right) +\vec{\nu}_e\right). 
\label{eq:uk}
\end{align}
%
This induces two types of the errors: first, the truncation
error due to the domain transformations $\mathcal{F}_{\langle *
\rangle}(\cdot)$; and second, the rounding error introduced by the
multiplication operation.  We represent these errors as non-deterministic
additive noise. More details can be found in Appendix~\ref{appendix:quantization}.

@@

Previous studies~\cite{gangli1} showed that the FWL affects the poles and
zeros positions, degrading the closed-loop dynamics, causing steady-state
errors and eventually de-stabilizing the system~\cite{Bessa16}.  However,
since the verifier also checks for the stability of the dynamics as a
precursor to safety, this case is handled inside the loop.  In the
following, we shall disregard these steady-state errors (caused by FWL
effects) when stability is ensured by synthesis, and then verify its safety
accounting for the finite-precision errors.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{Properties of a controllable model with observer}\label{sec:cof_verification}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%% In most real implementations, the internal state of a plant cannot be
%% directly measured but rather inferred from its
%% effect on the output. A plant model that minimises the difference
%% between the outputs is called an observer and it is a well known
%% paradigm in the control literature~\cite{astrom1997computer}.
